---
output:
  md_document:
    variant: markdown_github
bibliography: vignettes/references.bib
citation_package: natbib
biblio-style: apalike
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rgl.cry

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/rgl.cry)](https://CRAN.R-project.org/package=rgl.cry)
[![R-CMD-check](https://github.com/SaitouToshihide/rgl.cry/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/SaitouToshihide/rgl.cry/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

This is a use case for the ``cry`` and ``rgl`` packages.
This package provides tools for selected area electron diffraction (SAED) pattern and crystal structure visualization using the ``rgl`` and ``cry`` packages.
In particular, the ``cry_demo()`` and ``dp_demo()`` function read files in CIF (Crystallographic Information Framework) format and display SAED patterns and crystal structures.
The ``dp_demo()`` function also performs simple simulations of powder X-ray diffraction (PXRD) patterns, and the results can be saved to a file in the working directory.

The package has been tested on several platforms, including Linux on Crostini with a Core™ m3-8100Y Chromebook, I found that even on this low-powered platform, the performance was acceptable.


## Web Resources

https://saitoutoshihide.github.io/rgl.cry/


## Installation
```{r, eval=FALSE, echo=TRUE}
install.packages("rgl.cry")
```

## Example
A CIF file is read, and a reciprocal lattice map with a cell widget is drawn.
In this example, a file is not specified, so the system default is used.
```{r, eval=FALSE, echo=TRUE}
dp_demo()
```

A CIF file is read, and a crystal structure with a axis widget is drawn.
```{r, eval=FALSE, echo=TRUE}
cry_demo()
```

### Utility functions

The crystal and diffraction pattern are aligned and displayed.
```{r, eval=FALSE, echo=TRUE}
align("a")
align("ra")
align("30 30") # x, y (deg)
```

Select one or more atoms or reciprocal lattice points in the window.
The labels and Miller indices of the selected atoms or lattice points will be displayed.
```{r, eval=FALSE, echo=TRUE}
> dp_demo()
[1] 1
> select()
To select points, use dragging the left mouse button.
To finish, press ESC.
.
 [1] "1 1 -3" "1 0 -2" "2 0 -2" "1 1 -1" "1 0 0"  "2 0 0"  "1 1 1"  "1 0 2" 

> cry_demo()
[1] 2
> select()
To select points, use dragging the left mouse button.
To finish, press ESC.
.
[1] "Ti1" "Ti1"
```

### Extras
``dp_demo()`` can perform PXRD pattern simulation.
The result is saved as a file in the current directory by specifying options like this:
```{r, eval=FALSE, echo=TRUE}
## Output the simulation results of the PXDR pattern as a file.
dp_demo(xrd = TRUE)
```
The file looks like this:
```zsh
% sort -n +7 rgl.cry.dp.demo.2024-02-26_000000.dat
     h  k  l        d       absF         lp twotheta
137  0  0  0      Inf 308.117522        Inf  0.00000
136 -1  0  0 5.583222  17.389409 101.955123 15.87322
138  1  0  0 5.583222  17.389409 101.955123 15.87322
128  0 -2  0 4.453238  40.628676  63.820476 19.93782
146  0  2  0 4.453238  40.628676  63.820476 19.93782
129  1 -2  0 4.142661  17.444203  54.849532 21.44963
145 -1  2  0 4.142661  17.444203  54.849532 21.44963
```

## References

<div id="refs"></div>

## Acknowledgment

I would like to express my gratitude to the rgl, cry packages, and R for making this work possible.  The data for the scattering factors is based on the data from the KEK Report [@sc_table_kek].  For the coloring of the atoms, I used [@mol_color], and for the atomic radii, I used [@enwiki:1179864711].
